Using reference models in variable selection \vspace{.1in}

### 2020-04-23 paper/ref_approach.tex
added reference R package projpred
further references + language edits

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all variables.  The main assumption under any
reference model approach is that we operate in an
$\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is, we have
constructed a model which reflects our beliefs about
the future data in the best possible way and which has passed model checking and criticism
\citep[see, e.g.][]{gelman2013bayesian,gabry2019visualization}.  The reference model approach
has been used in Bayesian statistics at least since the seminal
work of \citet{paper:reference_lindley}. For more historical
references, see the papers by \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for most recent methodological
development see a paper by \citet{paper:projpred}.

Reference models have been also used in non-Bayesian contexts, where
\cite{harrell2015regression} describes them as full models that can be
thought of as a ``gold standard'' (for a given application).  For
example, \cite{faraggi2001understanding} deal with the necessity of
identifying interpretable risk groups in the context of survival data
using neural networks, which typically perform very well in terms of
prediction, but whose variables are difficult to be understood in
terms of relevance.  \cite{paul2008preconditioning}, using the term
preconditioning, explore approximating models fitting Lasso or
stepwise regression against consistent estimates $\hat{y}$ of a
reference model instead of the observed responses $y$.
Whatever the terminology or applied statistical framework, reference
models offer a powerful approach to improving variable selection as we
will demonstrate in the present paper.

@@

To motivate the further discussion and experiments, we start by simple
variable selection example using body fat data by \citet{johnson1996fitting}.
We compare the projective prediction approach (\emph{projpred},
\citet{paper:projpred}) which uses a reference model, and classic
stepwise backward regression (\emph{steplm}).
% \footnote{The analysis is based on \texttt{R}-notebook by Vehtari available at
%   \url{https://avehtari.github.io/modelselection/bodyfat.html}.}.
The experiments are implemented in \texttt{R} \citep{Rcore2018}.

The target variable of interest is the amount of body fat, which is
measured by immersing a person in water tank which is a complex and
expensive procedure. Additionally, we have information about 13
variables which are anthropometric measurements (e.g., height, weight
and circumference of different body parts). The
variables are highly correlated which causes additional
challenge in the variable selection. In total, we have 251
observations. The goal is to find the model which is able to predict
the amount of body fat well while requiring the least amount of
measurements for a new person.

@@

For the selection via projpred, the Bayesian reference model includes
all the variables using a regularised horseshoe prior~\citep{paper:rhs} on the
variable coefficients. Submodels are explored using
forward search (the results are not sensitive to whether forward or
backward search is used), and the predictive utility is the expected
log-predictive density (elpd) estimated using approximate leave-one-out cross-validation via Pareto-smoothed importance-sampling
\citep[PSIS-LOO-CV; ][]{paper:psis_loo}.  We select the smallest  submodel with an elpd score similar to the reference model
when taking into account the uncertainty in estimating the predictive
model performance. See Appendix A for brief review of projection
predictive approach, and papers by \citet{paper:model_selection} and
\citet{paper:projpred} for more details. The complete projpred approach is implemented in the \texttt{projpred} R package.

@@

We repeat the experiment with a modified data set by adding 84
unrelated noisy variables so that there are 100 variable in total. 
The last two rows of Table
\ref{tab:model_performances} show the cross-validated RMSE, the size
of the selected model and the number of selected noisy variables using
projpred or steplm. The results show that projpred has similar
predictive performance and the same number of selected variables as
with the original data, whereas the stepwise regression has worse
predictive performance and the number of selected variables is much
higher and include a large number of irrelevant variables.

@@

A properly designed reference model is able to filter parts of the
noise present in the data, and hence provide an improved and more
stable selection process. This holds even if the reference model does
not perfectly resemble the true data generating process. Moreover, our
analyses indicate that the substantial reduction of variance
attributable to noise is usually more important than small potential
bias due to model misspecification.  We argue that, regardless of how
the reference model is set up and used in the inference procedure, it
can be always seen as acting as a filter on the observed
data. Furthermore, regardless of what specific model selection method
is used, a reference model can be used instead of raw data during the
selection process to improve the stability and selection
performance. Our results indicate that the core reason why the
reference model based methods perform well is the reference model
itself, rather than the specific way of using it.  In general, the
less data we have and the more complex the estimated models, the
higher the benefit of using a reference models as the basis for
variable selection.

If one of the models to be compared is the full model which can be
used as a reference model, there is no additional cost of using a
reference model as it was estimated as part of the analysis anyway.
Sometimes including all the available variables in an
elaborate model can be computationally demanding.  In such a
case, even simpler screening or dimensionality reduction techniques,
as for example the supervised principal components
\citep{bair2006prediction,piironen2018} can produce useful reference
models.

@@

A good predictive model is able to filter part of the noise present in
the data. The noise is the main source of the instability and tends to
obscure the relevance of the variables in relation to the target
variable of interest. We demonstrate with the following simple
explanatory example taken from \cite{paper:projpred}. The data
generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim \N(0,1) && \nonumber \\ 
     Y|&f\sim \N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} \N(\sqrt{\rho}f,1-\rho) \quad &&j=1,\ldots,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} \N(0,1) &&j=k+1,\ldots,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy
observation. The first $k$ variables are strongly related to the
target variable $Y$ and correlated among themselves. Precisely, $\rho$
is the correlation among any pair of the first $k$ variables, whereas
$\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between
any relevant variable and, respectively, $f$ and $Y$. If we had an
infinite amount of observations, the sample correlation would be equal
to the true correlation between $X_j$ and $Y$. However, even in this
ideal asymptotic regime, this correlation would still remain a biased
indicator of the true relevance of each variable (represented by the
correlation between $X_j$ and $f$) due to the intrinsic noisy nature
of $Y$.

@@

Using the data generating mechanism \eqref{eq:simulated_data}, we
simulate data sets of different sizes with a relatively large number
of variables $p=70$ with $k=20$ of them being predictive.  We compare
the minimal subset variable selection perfomance of projection
predictive approach (with a reference model), a Bayesian stepwise
forward selection with and without a reference model, and maximum
likelihood stepwise forward selection with and without a reference
model:
\begin{itemize}
\item projpred: The projective prediction approach is used. 
  The reference model is a Bayesian linear regression
  model using the first five supervised principal components \citep{piironen2018} as predictors
  and the full posterior predictive distribution as the basis of the projection.   
  The search heuristic  is forward search and the predictive performance 
  is estimated via 10-fold cross-validation. The selection continues
  until the predictive performance is close to the predictive
  performance of the reference model. See more details in Appendix A.

\item Bayesian stepwise selection (without a reference model): At each step, the fitted model
 is a Bayesian linear regression using the regularised horseshoe
  prior and the variable excluded is the one with the highest Bayesian
  $p$-value defined as $\text{min}\{P(\theta\leq0 \; | \; D),P(\theta>0 \; | \; D)\}$.
The selection continues if the reduced model has an elpd score higher 
(i.e., better) than the current model.

@@

When the variable selection is repeated with different simulated data
sets, there is some variability in the selected variables. We
measure the stability of variable selection by computing the entropy
of observed distribution over different models. The smallest entropy
would be obtained if the approach always selected the same set of
variables, and the largest entropy would be observed if the approach
would always selected different sets of variables.
Figure \ref{fig:entropy} shows the entropy scores for the different compared
methods. Lower entropy corresponds to a more stable
selection. Highly correlated predictive variables may happen to be
selected alternately, thus making stability estimation of the
selection a non-trivial task. Entropy cannot distinguish the interchangeability 
due to correlation from instability. Thus, such a measure should be 
considerd as a relative and not as an absolute measure of stability.
The use of a reference model
improves the stability of steplm in variable selection slightly, while it makes
little difference for the Bayesian linear regression. The
projpred approach turns out to be far more stable than all
other methods. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

An alternative to minimal subset variable selection is \emph{complete
  variable selection} in which the goal is to find all relevant
variables that have some predictive information about the target. In
complete variable selection, it is possible that there are theoretically
relevant variables, but given finite noisy data we are not able infer their relevance.
The projection predictive approach was originally designed for the minimal subset
variable selection but we will test a simple iterative variant for the complete
variable selection case in the section.
 In addition, we analyse the benefits of the reference
model approach in combination with three other
methods which have been specifically designed for complete variable selection. 
As the criteria of selection performance, we evaluate the average false
discovery rate and the average
sensitivity (i.e., the ratio of the number of relevant selected
variables over the total number of relevant variables). We also provide
a comparison of the stability of the selection by means of a stability
measure proposed by \citet{paper:stability}, which goes from 0 to 1 and a higher value 
means a more stable selection.

@@

We modify the projection predictive approach for complete variable
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
variables, which yield to a model with a predictive performance
comparable to the full model's predictive performance. 
 The iterative projection repeats
the projpred selection for different iterations, at each time
excluding the variables selected in the previous
iterations from the search.
 At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size and ``best'' stands for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ is non-trivial, and we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive variables. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package, that is, $\alpha=0.16$.

@@

In the experiments shown in the next sections, we include an additional iterative
method which we refer to as `iterative lasso'. It consists of the same iterative
algorithm as iterative projpred expect for not using any reference model but
the lasso method for variable selection, instead.
That is, it uses the observed target values instead of predictions of the reference model.
Iterative lasso can help to distangle the effects
of the iterative procedure and the usage of a reference model in complete feature selection.

@@

We consider three alternative complete selection methods,
 the control of the local
false discovery rate \citep{paper:efron, efron2012large}, the
empirical Bayes median \citep{johnstone2004needles}, and the selection
by posterior credible intervals.

The control of the local false discovery rate consists of testing the
z-values $\{z_{j}\}_{j=1}^{p}$ on whether they belong to the
theoretical null distribution $f_{0}$ (i.e., the null hypothesis $H_0$
meaning no relevance) against the alternative hypothesis distribution
$f_{1}$. In our case $f_{0}$ corresponds to the standard normal
distribution (see expression \eqref{eq:normal_means_problem2}). The
quantity of interest is the local false discovery rate (loc.fdr)
defined as: \
\begin{equation}
\text{loc.fdr}(z)=P(H_{0}|z)=\frac{f_{0}(z)\pi_{0}}{f(z)},
\end{equation}
where $\pi_{0}$ is the prior probability of $H_0$ and
$f(z)=\pi_{0}f_{0}(z)+(1-\pi_{0})f_{1}(z)$ is the marginal
distribution of the z-values. The latter is estimated using splines
with 7 degrees of freedom. We select variables with local false
discovery rate below $0.2$, which is suggested by
\citet{efron2012large} as it corresponds to a Bayes factor larger than
36 (assuming $\pi_{0}\geq0.9$). The results of the comparison are not
sensitive to the specific value. To estimate $\pi_{0}$ from the data,
we use the default setting provided by the \texttt{R}-package
\texttt{locfdr} \citep{Efron+Turnbull+Narasimhan:2015:locfdr, Rlocfdr}.

@@

Figure \ref{fig:stability_iterated} shows the estimates of the stability
measure proposed by \cite{paper:stability} with 0.95 confidence
intervals based on 100 simulations. Such a measure takes in account
the variability of the subset of the selected variables at each
simulation (originally at each bootstrap sample), modelling the
selection of each variable as a Bernoulli process. Further details are
available in \cite{paper:stability}. The reference model helps in
improving the stability of the selection: Again, the benefits are
bigger when the problem is more difficult (small $n$ and $\rho$). In
addition, we observe less uncertainty in the stability estimates for
the reference approach (i.e., smaller width of the 95\% intervals),
which can be still connected to the overall stability of the
procedure. As in Figure \ref{fig:sensitivity_vs_fdr_iterated}, the iterative
projection does not perform well in the hardest scenarios.

@@

We conclude our complete selection experiments using the body fat
dataset one more time. As earlier we add noisy uncorrelated variables
to the original data to get a total of 100 variables.
Since we do not have a ground truth available with regard the original
variables of the data, we assume it is reasonable to consider all of
them relevant, at least to some degree. The artificially added
variables are naturally irrelevant by construction.
We compute
correlations between each variable and the target variable, that is
the amount of fat, and transform them by Fisher-Z-transformation. The
original assumption in order for \eqref{eq:fisher_transformation} to
hold is that the variables are jointly normally distributed. In our
experience the normal approximation in
\eqref{eq:fisher_transformation} is still reasonable, but after
rescaling by $\sqrt{n-3}$ we do not fix the variance to be one, and
instead estimate it from the data. We compare iteration projections, 
the control of the local false discovery rate (loc.fdr), 
the empirical Bayes median (EB.med),
and the selection by posterior credible intervals at level 90\%
(ci.90).  In order to vary the difficulty of the selection, we
bootstrap subsamples of different sizes, going from $n=50$ up to
$n=251$ (i.e., the full size of the data). For each condition, results
are averaged over 100 bootstrap samples of the respective size.

@@

In this example, we have used the reference model defined as a linear
regression over some supervised principal components, because it is
natural for a large number of correlating variables, has fairly good
predictive performance and is computationally efficient. We do not
argue that this is always the best choice and more sophisticated
models can lead to even better results. Here the purpose of the
experiments were to motivate the use of reference models in general,
and as we needed to average results over a lot of repetitions per
simulation condition, we preferred such a comparably simple and
computationally fast reference model.

@@

All Bayesian models used in the this paper have been implemented in
the probabilistic programming language \texttt{Stan}
\citep{paper:stan} and fit via dynamic Hamiltonian Monte Carlo
\citep{hoffman2014no,betancourt2017conceptual}. The code to run all
the experiments is available on GitHub
(\url{https://github.com/fpavone/ref-approach-paper}).

>>>>

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression \citep{venables2013modern,harrell2015regression}, the
observed empirical data distribution is utilised as a proxy of future
data usually in combination with cross-validation or information
criteria to provide estimates for out-of-sample predictive
performance.  In contrast, reference model based methods approximate
the future data generation mechanism using the predictive distribution
of a reference model, which can be, for example, a full-encompassing
model including all variables.  The main assumption under any
reference model approach is that we operate in an
$\mathcal{M}$-\textit{complete} framework
\citep{book:bernardo_smith,vehtari2012survey}, that is, we have
constructed a model which reflects our beliefs about
the future data in the best possible way and which has passed model checking and criticism
\citep[see, e.g.][]{gelman2013bayesian,gabry2019visualization}.  The reference model approach
has been used in Bayesian statistics at least since the seminal
work of \citet{paper:reference_lindley}. For more historical
references, see \citet{vehtari2012survey} and
\citet{paper:model_selection}, and for most recent methodological
developments see \citet{paper:projpred}.

Reference models have been also used in non-Bayesian contexts, in which
\cite{harrell2015regression} describes them as full models that can be
thought of as a ``gold standard'' (for a given application).  For
example, \cite{faraggi2001understanding} deal with the necessity of
identifying interpretable risk groups in the context of survival data
using neural networks, which typically perform very well in terms of
prediction, but whose variables are difficult to be understood in
terms of relevance.  \cite{paul2008preconditioning}, using the term
preconditioning, explore approximating models fitting Lasso or
stepwise regression against consistent estimates $\hat{y}$ of a
reference model instead of the observed responses $y$.
Whatever the terminology or applied statistical framework, reference
models offer a powerful approach to improving variable selection as we
will demonstrate in the present paper.

@@

To motivate the further discussion and experiments, we start by a simple
variable selection example using body fat data by \citet{johnson1996fitting}.
We compare the projective prediction approach (\emph{projpred},
\citet{paper:projpred}) which uses a reference model, and classic
stepwise backward regression (\emph{steplm}).
% \footnote{The analysis is based on \texttt{R}-notebook by Vehtari available at
%   \url{https://avehtari.github.io/modelselection/bodyfat.html}.}.
The experiments are implemented in \texttt{R} \citep{Rcore2018}.

The target variable of interest is the amount of body fat, which is
obtained by a complex and expensive procedure consisting in immersing a 
person in a water tank and carrying out different measuraments and computations. 
Additionally, we have information about 13
variables which are anthropometric measurements (e.g., height, weight
and circumference of different body parts). The
variables are highly correlated which causes additional
challenge in the variable selection. In total, we have 251
observations. The goal is to find the model which is able to predict
the amount of body fat well while requiring the least amount of
measurements for a new person.

@@

For the selection via projpred, the Bayesian reference model includes
all the variables using a regularised horseshoe prior~\citep{paper:rhs} on the
variable coefficients. Submodels are explored using
forward search (the results are not sensitive to whether forward or
backward search is used), and the predictive utility is the expected
log-predictive density (elpd) estimated using approximate leave-one-out cross-validation via Pareto-smoothed importance-sampling
\citep[PSIS-LOO-CV; ][]{paper:psis_loo}.  We select the smallest  submodel with an elpd score similar to the reference model
when taking into account the uncertainty in estimating the predictive
model performance. See Appendix A for brief review of projection
predictive approach, and papers by \citet{paper:model_selection} and
\citet{paper:projpred} for more details. The complete projpred approach is implemented in the \texttt{projpred} R package \citep{Rprojpred}.

@@

We repeat the experiment with a modified data set by adding 84
unrelated noisy variables so that there are 100 variables in total. 
The last two rows of Table
\ref{tab:model_performances} show the cross-validated RMSE, the size
of the selected model and the number of selected noisy variables using
projpred or steplm. The results show that projpred has similar
predictive performance and the same number of selected variables as
with the original data, whereas the stepwise regression has worse
predictive performance and the number of selected variables is much
higher and include a large number of irrelevant variables.

@@

A properly designed reference model is able to filter parts of the
noise present in the data, and hence to provide an improved and more
stable selection process. This holds even if the reference model does
not perfectly resemble the true data generating process. Moreover, our
analyses indicate that the substantial reduction of variance
attributable to noise is usually more important than small potential
bias due to model misspecification.  We argue that, regardless of how
the reference model is set up and used in the inference procedure, it
can be always seen as acting as a filter on the observed
data. Furthermore, regardless of what specific model selection method
is used, a reference model can be used instead of raw data during the
selection process to improve the stability and selection
performance. Our results indicate that the core reason why the
reference model based methods perform well is the reference model
itself, rather than the specific way of using it.  In general, the
less data we have and the more complex the estimated models are, the
higher is the benefit of using a reference models as the basis for
variable selection.

If one of the models to be compared is the full model, which can be
used as a reference model, there is no additional cost of using a
reference model as it was estimated as part of the analysis anyway.
Sometimes including all the available variables in an
elaborate model can be computationally demanding.  In such a
case, even simpler screening or dimensionality reduction techniques,
as for example the supervised principal components
\citep{bair2006prediction,piironen2018} can produce useful reference
models.

@@

A good predictive model is able to filter part of the noise present in
the data. The noise is the main source of the instability in the selection
and tends to
obscure the relevance of the variables in relation to the target
variable of interest. We demonstrate it with the following simple
explanatory example taken from \cite{paper:projpred}. The data
generation mechanism is
\begin{alignat}{2} \label{eq:simulated_data}
     &f\sim \N(0,1) && \nonumber \\ 
     Y|&f\sim \N(f,1) && \\
     X_{j}|&f \overset{iid}{\sim} \N(\sqrt{\rho}f,1-\rho) \quad &&j=1,\ldots,k \nonumber \\
     X_{j}|&f \overset{iid}{\sim} \N(0,1) &&j=k+1,\ldots,p \nonumber,
\end{alignat}
where $f$ is the latent variable of interest of which $Y$ is a noisy
observation. The first $k$ variables are strongly related to the
target variable $Y$ and correlated among themselves. Precisely, $\rho$
is the correlation among any pair of the first $k$ variables, whereas
$\sqrt{\rho}$ and $\sqrt{\rho/2}$ are the level of correlation between
any relevant variable and, respectively, $f$ and $Y$. If we had an
infinite amount of observations, the sample correlation would be equal
to the true correlation between $X_j$ and $Y$. However, even in this
ideal asymptotic regime, this correlation would still remain a biased
indicator of the true relevance of each variable (represented by the
correlation between $X_j$ and $f$) due to the intrinsic noisy nature
of $Y$.

@@

Using the data generating mechanism \eqref{eq:simulated_data}, we
simulate data sets of different sizes with a relatively large number
of variables $p=70$, with $k=20$ of them being predictive.  We compare
the minimal subset variable selection performance of the projection
predictive approach (which uses a reference model and it is referred to as projpred), 
a Bayesian stepwise
forward selection with and without a reference model, and maximum
likelihood stepwise forward selection with and without a reference
model (steplm). The following is a summary of the implementation of the compared methods:
\begin{itemize}
\item projpred: The projective prediction approach is used. 
  The reference model is a Bayesian linear regression
  model using the first five supervised principal components \citep{piironen2018} as predictors
  and the full posterior predictive distribution as the basis of the projection.   
  The search heuristic  is forward search and the predictive performance 
  is estimated via 10-fold cross-validation. The selection continues
  until the predictive performance is close to the predictive
  performance of the reference model. See more details in Appendix A.

\item Bayesian stepwise selection (without a reference model): At each step, the fitted model
 is a Bayesian linear regression using the regularised horseshoe
  prior and the variable excluded is the one with the highest Bayesian
  $p$-value defined as $\text{min}\{P(\theta\leq0 \; | \; D),P(\theta>0 \; | \; D)\}$,
  where $D$ stays for the observed data.
The selection continues if the reduced model has an elpd score higher 
(i.e., better) than the current model.

@@

When the variable selection is repeated with different simulated data
sets, there is some variability in the selected variables. We
measure the stability of variable selection by computing the entropy
of the observed distribution of the included variables over different models. The smallest entropy
would be obtained if the approach always selected the same set of
variables, and the largest entropy would be observed if the approach
would always selected different sets of variables. Therefore, lower entropy 
corresponds to a more stable selection. Highly correlated predictive variables may happen to be
selected alternately, thus making stability estimation of the
selection a non-trivial task. Entropy can not distinguish the interchangeability 
due to correlation from instability. Thus, such a measure should be 
considered as a relative, and not as an absolute, measure of stability.
Figure \ref{fig:entropy} shows the entropy scores for the different compared
methods. The use of a reference model
improves the stability of steplm in variable selection slightly, while it makes
little difference for the Bayesian linear regression. The
projpred approach turns out to be far more stable than all
other methods. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

An alternative to minimal subset variable selection is \emph{complete
  variable selection} in which the goal is to find all relevant
variables that have some predictive information about the target. In
complete variable selection, it is possible that there are theoretically
relevant variables, but given finite noisy data we are not able to infer their relevance.
The projection predictive approach was originally designed for the minimal subset
variable selection but we will test a simple iterative variant for the complete
variable selection case in this Section.
 In addition, we analyse the benefits of the reference
model approach in combination with three other
methods which have been specifically designed for complete variable selection. 
As the criteria of selection performance, we evaluate the average false
discovery rate and the average
sensitivity (i.e., the ratio of the number of relevant selected
variables over the total number of relevant variables). We also provide
a comparison of the stability of the selection by means of a stability
measure proposed by \citet{paper:stability}, which goes from 0 to 1 and a higher value 
means a more stable selection.

@@

We modify the projection predictive approach for complete variable
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
variables, which yield to a model with a predictive performance
comparable to the full model's predictive performance. 
 The iterative projection repeats
the projpred selection for different iterations, at each time
excluding the variables selected in the previous
iterations from the search.
 At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. This translates
in the following stopping rule at each iteration:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size and ``best'' stands for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ is non-trivial, and we have 
observed sensitivity of the selection to such a choice, mainly when using cross-validation with a small
number of observations or not very predictive variables. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package, that is, $\alpha=0.16$.

@@

In the experiments shown in the next sections, we include an additional iterative
method which we refer to as `iterative lasso'. It consists of the same iterative
algorithm as iterative projpred expect for not using any reference model, but
the lasso method for variable selection, instead.
That is, it uses the observed target values instead of predictions of the reference model.
The comparison with iterative lasso can help to disentangle the effects
of the iterative procedure and the usage of a reference model in complete feature selection.

@@

We consider three alternative complete selection methods:
 the control of the local
false discovery rate \citep{paper:efron, efron2012large}, the
empirical Bayes median \citep{johnstone2004needles}, and the selection
by posterior credible intervals.

The control of the local false discovery rate consists of testing the
z-values $\{z_{j}\}_{j=1}^{p}$ of a normal mean problem (explained in Section \ref{simulations}) 
on whether they belong to the
theoretical null distribution $f_{0}$ (i.e., the null hypothesis $H_0$
meaning no relevance) against the alternative hypothesis distribution
$f_{1}$. In our case $f_{0}$ corresponds to the standard normal
distribution (see expression \eqref{eq:normal_means_problem2}). The
quantity of interest is the local false discovery rate (loc.fdr)
defined as: \
\begin{equation}
\text{loc.fdr}(z)=P(H_{0}|z)=\frac{f_{0}(z)\pi_{0}}{f(z)},
\end{equation}
where $\pi_{0}$ is the prior probability of $H_0$ and
$f(z)=\pi_{0}f_{0}(z)+(1-\pi_{0})f_{1}(z)$ is the marginal
distribution of the z-values. The latter is estimated using splines
with 7 degrees of freedom. We select variables with local false
discovery rate below $0.2$, which is suggested by
\citet{efron2012large} as it corresponds to a Bayes factor larger than
36 (assuming $\pi_{0}\geq0.9$). The results of the comparison are not
sensitive to the specific value. To estimate $\pi_{0}$ from the data,
we use the default setting provided by the \texttt{R}-package
\texttt{locfdr} \citep{Efron+Turnbull+Narasimhan:2015:locfdr, Rlocfdr}.

@@

Figure \ref{fig:stability_iterated} shows the estimates of the stability
measure proposed by \cite{paper:stability} with 0.95 confidence
intervals based on 100 simulations. Such a measure takes in account
the variability of the subset of the selected variables at each
simulation (originally at each bootstrap sample), modelling the
selection of each variable as a Bernoulli process. Further details are
available in \cite{paper:stability}. The reference model helps in
improving the stability of the selection: again, the benefits are
bigger when the problem is more difficult (small $n$ and $\rho$). In
addition, we observe less uncertainty in the stability estimates for
the reference approach (i.e., smaller width of the 95\% intervals),
which can be still connected to the overall stability of the
procedure. As in Figure \ref{fig:sensitivity_vs_fdr_iterated}, the iterative
projection does not perform well in the hardest scenarios.

@@

We conclude our complete selection experiments using the body fat
dataset one more time. As earlier we add noisy uncorrelated variables
to the original data to get a total of 100 variables.
Since we do not have a ground truth available with regard the original
variables of the data, we assume it is reasonable to consider all of
them relevant, at least to some degree. The artificially added
variables are naturally irrelevant by construction.
We compute
correlations between each variable and the target variable, that is
the amount of fat, and transform them by Fisher-Z-transformation. The
original assumption in order for \eqref{eq:fisher_transformation} to
hold is that the variables are jointly normally distributed. In our
experience the normal approximation in
\eqref{eq:fisher_transformation} is still reasonable, but after
rescaling by $\sqrt{n-3}$ we do not fix the variance to be one, and
instead estimate it from the data. We compare the iterative projection, 
the control of the local false discovery rate (loc.fdr), 
the empirical Bayes median (EB.med)
and the selection by posterior credible intervals at level 90\%
(ci.90).  In order to vary the difficulty of the selection, we
bootstrap subsamples of different sizes, going from $n=50$ up to
$n=251$ (i.e., the full size of the data). For each condition, results
are averaged over 100 bootstrap samples of the respective size.

@@

In this example, we have used the reference model defined as a linear
regression over some supervised principal components, because it is
natural for a large number of correlating variables, and has fairly good
predictive performance plus it is computationally efficient. We do not
argue that this is always the best choice and more sophisticated
models can lead to even better results. Here the purpose of the
experiments were to motivate the use of reference models in general,
and as we needed to average results over a lot of repetitions per
simulation condition, we preferred such a comparably simple and
computationally fast reference model.

@@

All Bayesian models in this paper have been implemented in
the probabilistic programming language \texttt{Stan}
\citep{paper:stan} and fit via dynamic Hamiltonian Monte Carlo
\citep{hoffman2014no,betancourt2017conceptual}, through the R-packages 
\texttt{rstan} \citep{Rrstan} and \texttt{rstanarm} \citep{Rrstanarm}.
Graphics elaborations have been done using \texttt{ggplot2} \citep{Rggplot2}
and the \texttt{tidyverse} framework \citep{Rtidyverse}.
The code to run all
the experiments is available on GitHub
(\url{https://github.com/fpavone/ref-approach-paper}).
